Dust trapping in thin-ringed protoplanetary disks

### 2018-08-13 projects/dullemond/paper_dusttraps/ms.tex
New version of trapping paper.

%\documentclass[referee]{aa}
\documentclass{aa}
%\documentclass{mn2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\def\remark#1{{{\bf remark:} \bf #1}}
\def\action#1{{\bf #1}}
\def\putin#1{{\it #1}}
\def\revised#1{{\bf #1}}
\newcommand{\todo}[1]{\textcolor{red}{\bf #1}}
%
\def\aap{A\&A}
\def\apss{A\&SS}
\def\pasp{Proc.Astr.Soc.Pacific}
\def\pasj{Publ.Astr.Soc.Japan}
\def\araa{AnnRevA\& A}
\def\mnras{MNRAS}
\def\icarus{Icarus}
\def\nat{Nature}
\def\apj{ApJ}
\def\aj{AJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\jqsrt{JQSRT}
\def\max{\mathrm{max}}
\def\erg{\hbox{erg}}
\def\sec{\hbox{s}}
\def\Hz{\hbox{Hz}}
\def\gram{\hbox{g}}
\def\cm{\hbox{cm}}
\def\km{\hbox{km}}
\def\kel{\hbox{K}}
\def\KeV{\hbox{KeV}}
\def\MeV{\hbox{MeV}}
\def\eV{\hbox{eV}}
\def\yr{\hbox{yr}}
\def\AU{\hbox{AU}}
\def\ster{\hbox{ster}}
%\def\in{\mathrm{in}}
\def\out{\mathrm{out}}
\def\eff{\mathrm{eff}}
\def\rad{\mathrm{rad}}
\def\vrt{\mathrm{vert}}
\def\vset{\ensuremath{v_\mathrm{sett}}\xspace}
\def\fact{\ensuremath{\xi}\xspace}
\def\cs{\ensuremath{c_\mathrm{s}}\xspace}
\def\siggas{\ensuremath{\sigma_\mathrm{g}}\xspace}
\def\sigcoag{\ensuremath{\sigma_\mathrm{c}}\xspace}
\def\Tgas{\ensuremath{T_\mathrm{gas}}\xspace}
%\def\rhogas{\ensuremath{\rho_\mathrm{gas}}\xspace}
\def\rhogas{\ensuremath{\rho}\xspace}
\def\rhodust{\ensuremath{\rho_\mathrm{d}}\xspace}
\def\mugas{\ensuremath{\mu_\mathrm{gas}}\xspace}
\def\mp{\ensuremath{m_\mathrm{p}}\xspace}
\def\Hp{\ensuremath{H_\mathrm{p}}\xspace}
\def\Hs{\ensuremath{H_\mathrm{s}}\xspace}
\def\Omegak{\ensuremath{\Omega_\mathrm{K}}\xspace}
\def\Mstar{\ensuremath{M_\star}\xspace}
\def\Sc{\ensuremath{\mathrm{Sc}}\xspace}
\def\St{\ensuremath{\mathrm{St}}\xspace}
\def\Pt{\ensuremath{\mathrm{Pt}}\xspace}
\def\comma{\,,}
\def\fullstop{\,.}
%
\def\submitted{submitted\xspace}
\def\inprep{in prep.\xspace}
%
\begin{document}
%\thesaurus{02.01.2,08.03.4,08.06.2,08.16.5,13.09.6}
\title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
\authorrunning{Dullemond, Birnstiel, Troncoso, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, N.~Troncoso$^{3}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
\institute{
  (1) Zentrum f\"ur Astronomie, Heidelberg University, Albert Ueberle Str.~2, 69120 Heidelberg, Germany\\
  (2) {\bf XXXX}
  (3) {\bf XXXX}
  (4) {\bf XXXX}
  (5) {\bf XXXX}
  (6) {\bf XXXX}
} \date{\today}

@@

\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
\label{sec-data}
%
In this paper we focus on those sources of the DSHARP Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry about 3-D
line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.

\begin{table}
\begin{center}
\begin{tabular}{|c|ccc|}
\hline
\hline
Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
\hline
% AS 209     & 121              & 0.912        &  0.631 \\ % Data from 2018.05.13 Table
AS 209     & 121              & 0.832        &  1.413 \\ % 2018.07.02
Elias 24   & 136              & 0.776        &  6.03  \\ % 2018.07.02
% HD 163296  & 101              & 2.14         &  25.7  \\ % Data from 2018.05.13 Table
HD 163296  & 101              & 2.04         &  17.0  \\ % 2018.07.02
GW Lup     & 155              & 0.457        &  0.331 \\ % 2018.07.02
% HD 143006  & 166              & 0.813        &  0.871 \\ % Data estimated by Kees May/June 2018
HD 143006  & 165              & 1.78        &  3.80 \\  % 2018.07.02 ** Big change **
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-stellar-params}The stellar parameters assumed for the
  stars studied in this paper. Distance is in parsec and mass and luminosity
  are in units of the solar values.}
\end{table}

A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.
For an overview of the ALMA Large Programme we refer to {\bf Andrews et
  al.~XXXX}, and for an in-depth discussion on the data of the individual
sources we refer to {\bf [LIST HERE THE PAPERS]}.

\begin{figure*}
\centerline{\includegraphics[width=0.95\textwidth]{gallery_rings.pdf}}
\caption{\label{fig-obs-images}The continuum maps in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The eight highest contrast rings, which are the topic of this paper,
  are marked in the images. For a detailed description of these
  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
\end{figure*}

The high-contrast rings of these sources provide ``clean laboratories''
for testing the theory of dust trapping in a ring-by-ring manner.
Fig.\ \ref{fig-obs-profiles}
shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the five disks. The
procedure used to extract these radial profiles from the continuum maps
is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
HD 143006 we removed the blob]}. The
brightness is expressed as linear brightness temperature
$T_{\mathrm{br,lin}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
defined such that $B_\nu(T_{\mathrm{br,full}})=I_\nu$, with $B_\nu(T)$ the
Planck function, can have very different values, in particular at low
brightness. As a result $T_{\mathrm{br,full}}(r)$ tends to vastly
underestimate the contrast between the rings and the inter-ring
regions. Therefore, in this paper we plot the brightness using the linear
brightness temperature, which is proportional to the intensity in Jy/beam.}.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
\caption{\label{fig-obs-profiles}The linear brightness temperature in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The vertical axis is logarithmic to better show the contrast.
  The eight highest contrast rings are fitted by a Gaussian profile, shown
  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
  with the width of the ALMA beam. For a detailed description of these
  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
\end{figure*}

As is shown in Appendix \ref{sec-steady-state-analytic-trap-model}, for a
Gaussian pressure bump the solution to the radial dust mixing and drift problem
is, to first approximation, also a Gaussian, with a width smaller than, or equal
to, that of the gas pressure bump. Our analysis of the eight rings of this paper
therefore naturally starts with the fitting of the observed radial intensity
profiles with a Gaussian function. For some {\bf [CHECK]} of the rings of this
paper such Gaussian fits were done in the context of fitting the entire radial
intensity profile in the uv-plane to the ALMA data ({\bf Guzman et al.XXXX;
  More??}). Our fits may deviate slightly from those {\bf [CHECK]}, because here
we put emphasis on the shape of the peak of the rings, and deemphasize the
wings.

The fitting procedure is described in Appendix \ref{sec-gauss-fitting}.  The
fitting was done in the image plane, in order to make it possible to fit to a
limited radial range, i.e.\ selecting a single ring. The Gaussian fits are shown
as inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
Fig.~\ref{fig-obs-gaussfits} these Gaussian fits are shown in close-up. The
parameters of the best fits are listed in Table \ref{tab-gauss-params}.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
  paper. The blue curves are the observations, the green curves are the best fit
  Gaussian profiles. The ``fit range'' bar shows the radial range within which
  the Gauss was fitted to the data. The fit range was chosen to fit the part of
  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
  FWHM beam size of the observations. The grey band around the blue curve shows
  the estimated uncertainty of the data.}
\end{figure*}

@@

\begin{tabular}{|cc|cc|ccccc|cccccc|}

@@

Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\

@@

AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  15.8 &  10.9 &   0.6 &   0.41 &  33.3 &   1.81\\
AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  12.4 &   7.7 &   0.4 &   0.48 &  74.3 &   4.04\\
Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.36 &  41.1 &   2.23\\
HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  30.8 &  25.6 &   1.7 &   0.40 &  56.9 &   3.09\\
HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  25.3 &  20.2 &   0.8 &   0.28 &  48.3 &   2.62\\
GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &   0.30 &  36.3 &   1.97\\
HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  72.20 &   5.06 &   4.42 &  27.2 &  22.0 &   2.1 &   0.16 &  10.5 &   0.57\\
HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 112.51 &   7.88 &   7.49 &  21.6 &  16.6 &   2.0 &   0.16 &  26.5 &   1.44\\

@@

  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values
  $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$ are the inner and outer boundary of
  the fitting domain. $A$ is the peak linear brightness temperature of the
  best-fit Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$,
  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
  dust emission profile, also expressed as standard deviation in units of
  $\mathrm{au}$. The last six columns are derived from a simple disk model
  assumption: $T_{\mathrm{d}}$ is the midplane temperature of the disk (we
  assume gas and dust temperature to be equal) computed from
  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
  $\varphi=0.02$. From that one can derive the corresponding linear brightness
  temerature $T_{\mathrm{d,br,lin}}=B_\nu(T_{\mathrm{d}})c^2/2\nu^2$ where
  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of 0.13
  $\mathrm{cm}$ of ALMA band 6.  The fourth-to-last column expresses the
  deconvolved dust ring width in units of the disk pressure scale height $h_p$
  computed from $T_{\mathrm{d}}$. The $\tau$ column gives the estimated optical
  depth at the peak of the ring, calculated from
  $\tau=-\ln(1-A/T_{\mathrm{d,br,lin}})$.  The last two columns are the computed
  dust mass in the ring in units of Earth masses, assuming the small grain
  absorption opacity
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
  value
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). In making these
  estimates, the optically thin approximation was used. See Appendix
  \ref{sec-computing-dust-mass} for details. An estimate of the
  optical-depth-corrected dust mass can be made by multiplying the mass values
  in the table by $\tau/(1-e^{-\tau})$.}

@@

Given the closeness of the profiles to a Gaussian, we can use the Gaussian fit
to make estimates of the total mass of dust trapped in each of these rings. The
procedure is described in Appendix \ref{sec-computing-dust-mass}, and the
results, for the two opacity choices we make, are listed in Table
\ref{tab-gauss-params}. In making these estimates we have assumed the dust to be
made up of micron-sized dust, which is likely too small. The mass based on the
Ivezic opacity, $M_{\mathrm{d,iv}}$, will increase for $a>\lambda/2\pi$ by a
factor of $2\pi\,a/\lambda$ with $\lambda=0.13\,\mathrm{cm}$. For the mass based
on the Ricci opacity the factor first decreases when $a$ approaches the opacity
peak around $a=\lambda/2\pi$ and then increases for $a>\lambda/2\pi$. Both
behaviors can be seen as the inverse behavior of the opacities shown in
Fig.~\ref{fig-opacities}. For ever increasing grain size estimate, the mass
estimate will increase indefinitely for both opacity choices.

>>>>

%\documentclass[referee]{aa}
\documentclass{aa}
%\documentclass{mn2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\def\remark#1{{{\bf remark:} \bf #1}}
\def\action#1{{\bf #1}}
\def\putin#1{{\it #1}}
\def\revised#1{{\bf #1}}
\newcommand{\todo}[1]{\textcolor{red}{\bf #1}}
%
\def\aap{A\&A}
\def\apss{A\&SS}
\def\pasp{Proc.Astr.Soc.Pacific}
\def\pasj{Publ.Astr.Soc.Japan}
\def\araa{AnnRevA\& A}
\def\mnras{MNRAS}
\def\icarus{Icarus}
\def\nat{Nature}
\def\apj{ApJ}
\def\aj{AJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\jqsrt{JQSRT}
\def\max{\mathrm{max}}
\def\erg{\hbox{erg}}
\def\sec{\hbox{s}}
\def\Hz{\hbox{Hz}}
\def\gram{\hbox{g}}
\def\cm{\hbox{cm}}
\def\km{\hbox{km}}
\def\kel{\hbox{K}}
\def\KeV{\hbox{KeV}}
\def\MeV{\hbox{MeV}}
\def\eV{\hbox{eV}}
\def\yr{\hbox{yr}}
\def\AU{\hbox{AU}}
\def\ster{\hbox{ster}}
%\def\in{\mathrm{in}}
\def\out{\mathrm{out}}
\def\eff{\mathrm{eff}}
\def\rad{\mathrm{rad}}
\def\vrt{\mathrm{vert}}
\def\vset{\ensuremath{v_\mathrm{sett}}\xspace}
\def\fact{\ensuremath{\xi}\xspace}
\def\cs{\ensuremath{c_\mathrm{s}}\xspace}
\def\siggas{\ensuremath{\sigma_\mathrm{g}}\xspace}
\def\sigcoag{\ensuremath{\sigma_\mathrm{c}}\xspace}
\def\Tgas{\ensuremath{T_\mathrm{gas}}\xspace}
%\def\rhogas{\ensuremath{\rho_\mathrm{gas}}\xspace}
\def\rhogas{\ensuremath{\rho}\xspace}
\def\rhodust{\ensuremath{\rho_\mathrm{d}}\xspace}
\def\mugas{\ensuremath{\mu_\mathrm{gas}}\xspace}
\def\Hp{\ensuremath{H_\mathrm{p}}\xspace}
\def\Hs{\ensuremath{H_\mathrm{s}}\xspace}
\def\Omegak{\ensuremath{\Omega_\mathrm{K}}\xspace}
\def\Mstar{\ensuremath{M_\star}\xspace}
\def\Sc{\ensuremath{\mathrm{Sc}}\xspace}
\def\St{\ensuremath{\mathrm{St}}\xspace}
\def\Pt{\ensuremath{\mathrm{Pt}}\xspace}
\def\comma{\,,}
\def\fullstop{\,.}
%
\def\submitted{submitted\xspace}
\def\inprep{in prep.\xspace}
\def\paperdsharpandrews{Andrews et al.\ ({\bf 2018})}
\def\paperdsharphuangrings{Huang et al.\ ({\bf 2018})}
\def\paperdsharpisella{Isella et al.\ ({\bf 2018})}
\def\paperdsharpguzmann{Guzmann et al.\ ({\bf 2018})}
\def\paperdsharpbirnstiel{Birnstiel et al.\ ({\bf 2018})}
%
\begin{document}
%\thesaurus{02.01.2,08.03.4,08.06.2,08.16.5,13.09.6}
\title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
\authorrunning{Dullemond et al.}
\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, N.~Troncoso$^{3}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
\institute{
  (1) Zentrum f\"ur Astronomie, Heidelberg University, Albert Ueberle Str.~2, 69120 Heidelberg, Germany\\
  (2) {\bf XXXX}
  (3) {\bf XXXX}
  (4) {\bf XXXX}
  (5) {\bf XXXX}
  (6) {\bf XXXX}
} \date{\today}

@@

\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
\label{sec-data}
%
In this paper we focus on those sources of the DSHARP Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry much about 3-D
line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.

\begin{table}
\begin{center}
\begin{tabular}{|c|ccc|}
\hline
\hline
Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
\hline
AS 209     & 121              & 0.832        &  1.413 \\ % 2018.07.02
Elias 24   & 136              & 0.776        &  6.03  \\ % 2018.07.02
HD 163296  & 101              & 2.04         &  17.0  \\ % 2018.07.02
GW Lup     & 155              & 0.457        &  0.331 \\ % 2018.07.02
HD 143006  & 165              & 1.78         &  3.80 \\  % 2018.07.02 ** Big change **
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-stellar-params}The stellar parameters assumed for the stars
  studied in this paper. Distance is in parsec and mass and luminosity are in
  units of the solar values. More details, as well as references and uncertainty
  estimates, can be found in \paperdsharpandrews{}.}
\end{table}

A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.  For an
overview of the ALMA Large Programme we refer to \paperdsharpandrews{}, and for
an in-depth discussion on the data of the individual sources we refer to
\paperdsharphuangrings{}, \paperdsharpisella{}, \paperdsharpguzmann{}.

\begin{figure*}
\centerline{\includegraphics[width=0.95\textwidth]{gallery_rings.pdf}}
\caption{\label{fig-obs-images}The continuum maps in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The eight highest contrast rings, which are the topic of this paper,
  are marked in the images. For a detailed description of these
  data, see \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
  for HD 163296, and \paperdsharphuangrings{} for the rest.}
\end{figure*}

The high-contrast rings of these sources provide ``clean laboratories'' for
testing the theory of dust trapping in a ring-by-ring manner.
Fig.\ \ref{fig-obs-profiles} shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the five disks. The
procedure used to extract these radial profiles from the continuum maps is
described by \paperdsharphuangrings{}, though we removed the blob seen in HD
163296 and HD 143006 for this work.

The brightness profiles are expressed as linear brightness temperature
$I_\nu^{\mathrm{lb}}$, which we define in Appendix \ref{sec-conventions}. The
choice of linear brightness temperature over full brightness temperature is
important, because the full brightness temperature tends to vastly underestimate
the contrast between the rings and the inter-ring regions. The linear brightness
temperature is instead linearly proportional to the intensity in Jy/beam, and
thus represents the contrast appropriately.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
\caption{\label{fig-obs-profiles}The linear brightness temperature in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The vertical axis is logarithmic to better show the contrast.
  The eight highest contrast rings are fitted by a Gaussian profile, shown
  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
  with the width of the ALMA beam. For a detailed description of these
  data, \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
  for HD 163296, and \paperdsharphuangrings{} for the rest.}
\end{figure*}

\section{Fitting a Gaussian profile to the ring emission}
\label{sec-gauss-fits}
%
As we will discuss later (Section \ref{sec-rings-as-dust-traps} and Appendix
\ref{sec-steady-state-analytic-trap-model}), for a
Gaussian pressure bump the solution to the radial dust mixing and drift problem
is, to first approximation, also a Gaussian, with a width smaller than, or equal
to, that of the gas pressure bump. Our analysis of the eight rings of this paper
therefore naturally starts with the fitting of the observed radial intensity
profiles with a Gaussian function. For some {\bf [CHECK]} of the rings of this
paper such Gaussian fits were done in the context of fitting the entire radial
intensity profile in the uv-plane to the ALMA data ({\bf Guzman et al.XXXX;
  More??}). Our fits may deviate slightly from those {\bf [CHECK]}, because here
we put emphasis on the shape of the peak of the rings, and deemphasize the
wings. {\bf CHECK}.

The aim is now to find, for each ring, a Gaussian linear brightness temperature
profile
\begin{equation}\label{eq-gauss-lin-br-temp}
I_{\nu}^{\mathrm{lb,gauss}}(r) = A\,\exp\left(-\frac{(r-r_0)^2}{2\sigma^2}\right)
\end{equation}
that best describes the ring. To be more precise: We determine the values of
$A$, $r_0$ and $\sigma$ for which Eq.~(\ref{eq-gauss-lin-br-temp}) best fits the
observed linear brightness temperature profile $I_{\nu}^{\mathrm{lb,obs}}(r)$
shown in Fig.~\ref{fig-obs-profiles} within a prescribed radial domain as given
in Table \ref{tab-gauss-params}. Details of the fitting procedure are described
in Appendix \ref{sec-gauss-fitting-procedure}. The Gaussian fits are shown as
inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
Fig.~\ref{fig-obs-gaussfits} these Gaussian fits are shown in close-up. The
parameters of the best fits are listed in Table \ref{tab-gauss-params}.  For the
error estimates on these values, see Appendix \ref{sec-gauss-fitting-procedure}.

{\bf [ADD APPENDIX WITH DETAILS OF FITS AND ERROR ESTIMATES]}

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
  paper. The blue curves are the observations, the orange curves are the best fit
  Gaussian profiles. The ``fit range'' bar shows the radial range within which
  the Gauss was fitted to the data. The fit range was chosen to fit the part of
  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
  FWHM beam size of the observations. The grey band around the blue curve shows
  the estimated uncertainty of the data.}
\end{figure*}

@@

\begin{tabular}{|cc|cc|cccccc|ccccc|}

@@

%Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
%            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
%Source & Ring & beam & domain & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
%            & & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
Source & Ring & beam & domain & $A$ & $A_{\mathrm{dec}}$ & $r_0$ & $\mathrm{FWHM}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $B_{\mathrm{d}\nu}^{\mathrm{lb}}$ & $w_d/h_p$ & $\tau_\nu^{\mathrm{peak}}$ & $M_{\mathrm{d}}$\\
            & & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$\\

@@

% Note: New table with new beam size estimates 2018.08.06
% Note: Added beam size to table, changed a bit the layout, and used new opacities for mass estimates (though still needs improving) 2018.08.07
% Note: Removed the Ivezic mass column
% Note: Added A_dec
AS 209     & 1 & 37 &  69 --  79 &   3.64 &   4.18 &  74.1 &  75.4 &   3.87 &   3.38 &  15.8 &  10.9 &   0.6 &   0.48 &  28.5\\
AS 209     & 2 & 37 & 115 -- 125 &   2.95 &   3.23 & 120.3 &  90.3 &   4.64 &   4.23 &  12.4 &   7.7 &   0.4 &   0.55 &  60.7\\
Elias 24   & 1 & 35 &  72 --  82 &   5.22 &   5.69 &  76.9 &  88.0 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.40 &  33.4\\
HD 163296  & 1 & 43 &  60 --  76 &   8.48 &   8.76 &  67.4 & 171.7 &   7.36 &   7.13 &  30.8 &  25.6 &   1.6 &   0.42 &  43.8\\
HD 163296  & 2 & 43 &  94 -- 104 &   4.92 &   5.20 &  99.9 & 133.6 &   5.73 &   5.42 &  25.3 &  20.2 &   0.8 &   0.30 &  38.0\\
GW Lup     & 1 & 44 &  79 --  89 &   1.48 &   1.87 &  85.1 &  72.1 &   4.74 &   3.76 &  10.3 &   5.7 &   0.5 &   0.40 &  34.1\\
HD 143006  & 1 & 45 &  35 --  45 &   3.21 &   4.11 &  41.0 &  72.2 &   5.06 &   3.95 &  27.2 &  22.0 &   1.9 &   0.21 &  10.0\\
HD 143006  & 2 & 45 &  59 --  72 &   2.48 &   2.71 &  64.8 & 112.5 &   7.88 &   7.22 &  21.6 &  16.6 &   2.0 &   0.18 &  21.6\\

@@

  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The ``beam''
  column gives the beam size (full-width-at-half-max). Given that in reality the
  beam is elliptic, and the disk has an inclination, this is merely an estimate.
  The ``domain'' column gives the radial domain within which the Gaussian fit
  was made. $A$ is the peak linear brightness temperature of the best-fit
  Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$, $\mathrm{FWHM}$ is
  the width expressed as full-width-at-half-maximum in units of milliarcsecond,
  $\sigma$ is its width expressed as standard deviation in units of
  $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved) dust
  emission profile, also expressed as standard deviation in units of
  $\mathrm{au}$. The $A_{\mathrm{dec}}$ column is like $A$, but corrected for
  the deconvolution (see Eq.~\ref{eq-a-deconv}). The last five columns are
  derived from a simple disk model assumption: $T_{\mathrm{d}}$ is the midplane
  temperature of the disk (we assume gas and dust temperature to be equal)
  computed from Eq.~(\ref{eq-disk-temperature-model}), assuming a flaring angle
  of $\varphi=0.02$. From that one can derive the corresponding linear
  brightness temperature $B_{\mathrm{d}\nu}^{\mathrm{lb}}$ using
  Eq.~(\ref{eq-app-bnu-bright}) for $\nu=c/(0.13\,\mathrm{cm})$ (ALMA band 6).
  The $w_d/h_p$ column expresses the deconvolved dust ring width in units of the
  disk pressure scale height $h_p$ computed from $T_{\mathrm{d}}$. The
  $\tau_\nu^{\mathrm{peak}}$ column gives the estimated optical depth at the peak of the ring,
  calculated from Eq.~(\ref{eq-tau-estimate}). The last column is the computed dust mass in the
  ring in units of Earth masses. See Appendix \ref{sec-computing-dust-mass} for
  details. In making these mass estimates we use the DSHARP dust opacity model
  (\paperdsharpbirnstiel{}) for a grain radius of $a=0.1\,\mathrm{cm}$, which
  yields an absorption opacity
  $\kappa_\nu^{\mathrm{abs}}(\lambda=0.13\,\mathrm{cm})=1.93\,\mathrm{cm}^2/\mathrm{g}$.
  The uncertainty of the opacity value is, however, very large, since we do not
  know apriori the grain size distribution.}

@@

\section{Optically thin analysis}
\label{sec-opt-thin-analysis}
%
As a first guess, we will assume that we can ignore optical depth effects,
i.e.\ that the thermal emission of the dust is optically thin. The linear
brightness temperature profiles shown in Section \ref{sec-data}, after
deconvolution with the beam, are then linear maps of the spatial distribution of
dust, if we ignore any temperature gradients across these rings.  In other
words: the deconvolved versions of the Gaussian fits of Section
\ref{sec-gauss-fits} (Eq.~\ref{eq-gauss-lin-br-temp-deconv}) will then directly
translate in Gaussian models of the radial dust distribution.

The conversion between the deconvolved observed linear brightness temperature
profile $I_\nu^{\mathrm{lb,dec}}(r)$ and the dust surface density profile
$\Sigma_d(r)$ is then
\begin{equation}\label{eq-optthin-conversion-linbright-sigmad}
  %I_\nu^{\mathrm{lb}} = \Sigma_d\,\kappa_\nu^{\mathrm{abs}}\, B_\nu^{\mathrm{lb}}(T_d)
  \Sigma_d(r) =  \frac{I_\nu^{\mathrm{lb,dec}}(r)}{\kappa_\nu^{\mathrm{abs}}\, B_\nu^{\mathrm{lb}}(T_d)} 
\end{equation}
where $T_d$ is the temperature of the dust, $\kappa_\nu^{\mathrm{abs}}$ is the
absorption opacity, and $B_\nu^{\mathrm{lb}}(T_d)$ is the Planck function
expressed as linear brightness temperature (Eq.~\ref{eq-app-bnu-bright}).
